### 2019-03-21 1_introduction.tex
Update graphs for evaluation.
Rewrite introduction.

Challenges in heterogeneous and extreme-scale parallel and distributed
computing have lead to the rise of a wide variety of new programming
models and runtime systems. However, efforts to understand the
performance behavior of these system have been limited. In many cases,
programming models are evaluated with weak and/or strong scaling of a
bespoke set of carefully hand-written applications, which make it
difficult to understand how the results generalize to other
applications, and impossible to compare across systems. Furthermore,
these are not limit studies, and thus do nothing to elucide the
boundaries of the performance envelops of the respective systems.

We introduce \emph{minimum effective task granularity} (METG), a
metric of application performance specifically designed to explore the
fundamental performance behavior of parallel and distributed runtime
systems. METG can be used to identify the baseline overhead, limit of
scalability, and ability to hide and overlap communication of such
systems. A key insight of the METG metric is that large \emph{tasks},
or units of work, can be used to hide almost any amount of
overhead. Thus METG considers the \emph{minimum} task granularity that
can be executed. This is a lower bound, so as long as the average task
granularity of a given application exceeds this value, the runtime
system should be able to execute it efficiently.

A variety of runtime systems are able to make use of parallelism
internally for a variety of purposes, from analysis of tasks, to
scheduling or the orchestration of data movement and
synchronization. In order to mitigate unwanted influence due to
unreasonable allocation of resources to such internal uses, METG
concerns \emph{effective} task execution, where the overall
computational efficiency meets some threshold (say, 50\%). If the
METG(50\%) of an application is, say, 1~ms, then this indicates that
the application can be executed with tasks as small as 1~ms while
maintaining at least 50\% overall efficiency.

We present Task Bench, a benchmark designed specifically to study the
performance behavior of parallel and distributed runtime systems at
the limits of runtime overhead. Task Bench is a \emph{parameterized}
benchmark, characterized by an \emph{iteration space} of tasks (with
one task per point in the iteration space) and a \emph{dependence
  relation} which determines how tasks depend on other tasks. By
formulating the benchmark in this way, we make it possible to rapidly
simulate a wide variety of communication patterns relevant in
scientific computing, such as trivial parallelism, halo exchanges (for
structured and unstructured mesh-based computations), sweeps (such as
for the discrete ordinates method), FFTs, trees (for divide and
conquer algorithms), and so on. Task Bench can be configured to
execute a variety of kernels, such as compute- and memory-bound
kernels, and load imbalanced variants of the above. Task Bench can
also be used to stress communication-bound cases, by configuring the
amount of data that is produced by each task.

In order to explore the impact of task parallelism on runtime
performance, Task Bench can also execute multiple task graphs
simultaneously. Runtime systems which support asynchronous execution
can use this additional scheduling flexibility to hide communication
overheads or to mitigate load imbalance. METG makes it possible to
quantify the degree to which various systems are successful at doing
so, and to measure if (and when) such factors are more, or less,
important than baseline runtime overhead.

We provide a comprehensive evaluation of Task Bench implementations in
13 different runtime systems, including Chapel~\cite{Chapel07},
Charm++~\cite{Charmpp93}, MPI~\cite{MPI}, OmpSs~\cite{OmpSs11},
OpenMP~\cite{OpenMPSpec40}, PaRSEC~\cite{PARSEC13},
Realm~\cite{Realm14}, Regent~\cite{Regent15}, Spark~\cite{Spark10},
StarPU~\cite{StarPU11}, Swift/T~\cite{Wozniak13},
TensorFlow~\cite{TensorFlow15}, and X10~\cite{X1005}. Using the METG
metric, we determine the baseline runtime overhead of each system
along with overheads associated with increasing number and/or
variations in the pattern of dependencies. We also explore each
system's ability to scale, hide communication overhead, and mitigate
load imbalance.

@@

\begin{itemize}
\item Section~\ref{sec:metg} defines the METG metric and its
  relationship to quantities of interest to application developers.
\item Section~\ref{sec:task-bench} describes the design of Task Bench.
\item Section~\ref{sec:implementation} details the implementation of
  Task Bench in 13 runtime systems.
\item Section~\ref{sec:evaluation} provides a comprehensive evaluation
  of Task Bench performance on up to 256 nodes of the Cori
  supercomputer.
\end{itemize}

>>>>

Challenges in heterogeneous and extreme-scale parallel and distributed
computing have lead to the rise of a wide variety of new programming
models, languages and runtime systems. These systems often provide
capabilities such as asynchronous execution of units of work, called
\emph{tasks}, coordinated in some manner with data movement and
synchronization. But while the features of these systems are widely
written about in the literature, a large and often difficult to answer
question is how different systems will perform for a given
application.

One approach to answering this question has focused around the
development of \emph{proxy-} or \emph{mini-apps}. So called because
they embody the logic of a larger application, but in reduced form,
mini-apps are popular because they can provide insight without
incurring the costly expense of developing a full-blown application
code. However, despite the name, mini-apps still requite significant
investment to develop, especially to a level of quality need to reach
a publishable state. In many cases, the effort to tune for performance
exceeds the effort to develop the mini-app in the first place. This
can be exacerbated if the system in question is undergoing concurrent
development. As a result, implementations of mini-apps, though
well-intentioned, often do not reach a level of maturity required to
perform a broad comparison between systems, and as a result there are
very few comparative studies involving multiple programming systems
that reach publication.

A compounding issue is one of methodology: how does one predict how
the performance observed for a mini-app will generalize to a full
application, even one closely related? Part of the problem here is
that weak and strong scaling, which are by far the most commonly
reported metrics for performance of both mini- and full applications,
do nothing to elucidate the performance envelop of a particular
system. Weak scaling is notorious for being able to hide arbitrary
amounts of overhead, as long as the problem size is sufficiently
large. And strong scaling runs the risk of entangling unrelated
effects: namely, the baseline overhead of a given system, with the
growth of communication latencies, and potentially the saturation of
global resources such as communication and I/O bandwidth, when using
progressively larger portions of the machine.

To address the first of these issues, we present Task Bench, a
\emph{parameterized} benchmark designed for exploring the performance
behaviors of programming systems under a variety of conditions. A
critical property of Task Bench is that, while powerful, it is simple
enough to be implemented in a wide variety of systems with reasonable
effort, to a level of quality necessary for comparative study.

Task Bench consists of a configurable \emph{task graph} with tasks for
each point in an iteration space. Tasks are connected by dependencies
as determined by a \emph{dependence relation}. This design permits a
very concise description of a wide variety of patterns relevant to
scientific computing: trivial parallelism, halo exchanges (such as
seen in structured and unstructured mesh codes), sweeps (such as used
in the discrete ordinates method of radiation simulation), FFTs, trees
(for divide and conquer algorithms), and so on. The tasks at each
point in the task graph can be configured to execute kernels with a
variety of computational properties, including compute- and
memory-bound operations of varying durations. The dependencies between
tasks can be configured to carry payloads of varying size, permitting
the design of communication-bound cases. Finally, multiple
(potentially heterogeneous) task graphs can be executed concurrently
to introduce task parallelism into the workload. Together, these
design elements permit the exploration of a broad space of application
performance behaviors.

To resolve the methodological issue, we introduce a new metric called
\emph{minimum effective task granularity} (METG). Intuitively, METG is
the smallest task granularity which can be executed while maintaining
a given level of computational efficiency (e.g., as measured by the
percentage of peak FLOPS achieved). METG captures the essence of what
is important in a weak or strong scaling study: that is, the behavior
at the limit of scalability. For weak scaling, METG corresponds
directly to the smallest problem size that can be weak scaled with a
given level of efficiency. For strong scaling, METG can be used to
compute the point at which strong scaling can be expected to tail
off. In either case, METG is a useful, intuitive measure as it
encapsulates the task granularity at which systems can be relied on to
execute efficiently. As a rule of thumb, as long as an application's
average task granularity exceeds the METG, the application should
execute efficiently.

Together, METG and Task Bench help to mitigate issues which are common
in limit studies on the task scheduling throughput of runtime
systems. Such studies typically employ a metric called \emph{tasks per
  second} (TPS), which is almost universally measured with trivial
dependencies and empty tasks. While TPS is certainly an upper bound on
task scheduling throughput, it is not a tight bound, as the cost of
non-trivial dependencies can be significant and sometimes adversely
impact scalability. This issue is easily fixed by running non-trivial
configurations of Task Bench.

But there is a deeper issue with the use of empty tasks. In all
realistic scenarios, runtime systems must share resources with the
application, either explicitly by e.g., allocating cores for runtime
use, or implicitly by sharing a time slice on the same cores used for
application tasks, and other resources such as memory bandwidth and
processor cache capacity. The use of empty tasks makes the application
share of these resources appear to be zero, which permits usage by the
runtime system to be unconstrained. How then is anyone supposed to
determine what resource usage to consider reasonable on the part of
the runtime system? Only by requiring the application usage of
resources to meet a certain minimum level, as in METG, can we ensure
that runtime system resource usage is constrained to realistic levels.

We conduct a comprehensive comparative study of Task Bench
implementations in 13 different programming systems, including
Chapel~\cite{Chapel07}, Charm++~\cite{Charmpp93}, MPI~\cite{MPI},
OmpSs~\cite{OmpSs11}, OpenMP~\cite{OpenMPSpec40},
PaRSEC~\cite{PARSEC13}, Realm~\cite{Realm14}, Regent~\cite{Regent15},
Spark~\cite{Spark10}, StarPU~\cite{StarPU11},
Swift/T~\cite{Wozniak13}, TensorFlow~\cite{TensorFlow15}, and
X10~\cite{X1005}. Using the METG metric, we determine the baseline
runtime overhead of each system along with overheads associated with
increasing number and/or variations in the pattern of dependencies. We
also explore each system's ability to scale, hide communication
overhead, and mitigate load imbalance.

@@

\begin{itemize}
\item Section~\ref{sec:metg} defines the METG metric and its
  relationship to quantities of interest to application developers.
\item Section~\ref{sec:task-bench} describes the design of Task Bench.
\item Section~\ref{sec:implementation} details the implementation of
  Task Bench in 13 programming systems.
\item Section~\ref{sec:evaluation} provides a comprehensive evaluation
  of Task Bench performance on up to 256 Haswell nodes of the Cori
  supercomputer.
\end{itemize}
